---
title: "Exp042 Ploidy Quantification"
author: "Alex Lessenger"
date: "2024-03-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# To do:
2. Compare different int rings
3. Check whether averaging C values per worm or per ring decreases noise

```{r}
library(tidyverse)
library(ggbeeswarm)
library(extrafont)
loadfonts('pdf', quiet = T)

atltheme = function() {
  theme(legend.position = 'none',
        panel.background = element_blank(),
        strip.background = element_blank(), 
        legend.key = element_blank(),
        axis.line = element_line(color = "black", linewidth = 0.3),
        text = element_text(family = 'Arial', color = "black", size = 8),
        axis.text = element_text(family = 'Arial', color = "black", size = 8))
}
```

```{r}
raw_ploidy <- read.csv('quantification_data/Exp042_raw_ploidy.csv', stringsAsFactors = F)
raw_ploidy_expanded <- read.csv('quantification_data/Exp042_raw_ploidy_expanded1.csv', stringsAsFactors = F)
```

## Ploidy measurements V1: normalize to tetraploid pachytene germline
```{r}
bgcor_ploidy <- raw_ploidy %>%
  mutate(genotype = str_extract(image, '(?<=240129_)(\\w+)(?=_)'),
         cell_type = case_when(tissue == 'germline' ~ 'germ',
                               sub_tissue %in% c(1,8,9) ~ 'int1,8,9',
                               sub_tissue == 2 ~ 'int2',
                               TRUE ~ 'int3-7'),
         bckg = 123, # from manually-measured cytoplasm of unstained worms
         bgcor_mean_intensity = mean - bckg,
         bgcor_total_intensity = bgcor_mean_intensity * number_of_pixels)

ploidy <- bgcor_ploidy %>%
  filter(tissue == 'germline', number_of_pixels_on_border == 0) %>%
  group_by(image) %>%
  summarize(haploid_intensity = mean(bgcor_total_intensity) / 4) %>%
  right_join(bgcor_ploidy) %>%
  mutate(c_val = bgcor_total_intensity / haploid_intensity) %>%
  filter(tissue == 'intestine',
         number_of_pixels_on_border < 0.1 * number_of_pixels)
```

```{r}
p <- ploidy %>%
  ggplot(aes(c_val, fill = genotype)) +
  scale_x_log10(breaks = c(1,2,4,8,16,32,64)) +
  geom_histogram(alpha = 0.25, bins = 30, position = 'identity')
binwidth = layer_data(p) %>% mutate(w=xmax-xmin) %>% pull(w) %>% median
p +
  stat_bin(aes(color = genotype), geom = 'step',
           binwidth = binwidth, position=position_nudge(x=-0.5*binwidth)) +
  scale_fill_manual(values = c('firebrick2', 'gray50', 'dodgerblue2')) +
  scale_color_manual(values = c('firebrick3','gray20', 'dodgerblue3')) +
  annotation_logticks(sides = 'b') +
  atltheme() 
```

## Ploidy measurements V2: normalize to expanded labels of tetraploid pachytene germline
V1 Ploidy measurements gives fractional ploidies that are slightly higher than expected 2^n values, suggesting that the germline normalization value is slightly too low. This could be caused by under-segmented cells (which were never manually corrected). To fix this, I used `skimage.segmentation.expand_labels()` to expand the labels of all nuclei. Here, I normalize the non-expanded (but manually corrected) intestine values to the expanded germline values
```{r}
bgcor_ploidy_expanded <- raw_ploidy_expanded %>%
  mutate(genotype = str_extract(image, '(?<=240129_)(\\w+)(?=_)'),
         cell_type = case_when(tissue == 'germline' ~ 'germ',
                               sub_tissue %in% c(1,8,9) ~ 'int1,8,9',
                               sub_tissue == 2 ~ 'int2',
                               TRUE ~ 'int3-7'),
         bckg = 123, # from manually-measured cytoplasm of unstained worms
         bgcor_mean_intensity = mean - bckg,
         bgcor_total_intensity = bgcor_mean_intensity * number_of_pixels)

ploidy2 <- bgcor_ploidy_expanded %>%
  filter(tissue == 'germline', number_of_pixels_on_border <= 0.0*number_of_pixels) %>%
  group_by(image) %>%
  summarize(haploid_intensity = mean(bgcor_total_intensity) / 4) %>%
  right_join(bgcor_ploidy) %>%
  mutate(c_val = bgcor_total_intensity / haploid_intensity) %>%
  filter(tissue == 'intestine',
         number_of_pixels_on_border <= 0.05*number_of_pixels,
         c_val > 1.8)

p <- ploidy2 %>%
  ggplot(aes(c_val, fill = genotype)) +
  scale_x_log10(breaks = c(1,2,4,8,16,32,64)) +
  geom_histogram(alpha = 0.5, bins = 30, position = 'identity')
binwidth = layer_data(p) %>% mutate(w=xmax-xmin) %>% pull(w) %>% median
p1 <- p +
  stat_bin(aes(color = genotype), geom = 'step',
           binwidth = binwidth, position=position_nudge(x=-0.5*binwidth)) +
  # scale_fill_manual(values = c('firebrick2', 'black', 'dodgerblue2')) +
  scale_fill_manual(values = c('#FF4F57','grey30', '#00328A')) +
  scale_color_manual(values = c('#FF4F57','black', '#00328A')) +
  annotation_logticks(sides = 'b') +
  labs(x = 'nucleus ploidy (C)') +
  atltheme() 
# ggsave('R_plots/Exp042 ploidy histogram V2.pdf', p1, device = cairo_pdf, units = 'in', width = 3.2, height = 1.9)
# ggsave('R_plots/Exp042 ploidy histogram V2.png', p1, device = 'png', units = 'in', width = 3.2, height = 1.9, dpi = 300)
```

```{r plot without F2s}
p <- ploidy2 %>%
  filter(genotype != 'balminF2') %>%
  ggplot(aes(c_val, fill = genotype)) +
  scale_x_log10(breaks = c(1,2,4,8,16,32,64)) +
  geom_histogram(alpha = 0.5, bins = 30, position = 'identity')
binwidth = layer_data(p) %>% mutate(w=xmax-xmin) %>% pull(w) %>% median
p1 <- p +
  stat_bin(aes(color = genotype), geom = 'step',
           binwidth = binwidth, position=position_nudge(x=-0.5*binwidth)) +
  # scale_fill_manual(values = c('firebrick2', 'black', 'dodgerblue2')) +
  scale_fill_manual(values = c('#FF4F57', '#00328A')) +
  scale_color_manual(values = c('#FF4F57', '#00328A')) +
  annotation_logticks(sides = 'b') +
  labs(x = 'nucleus ploidy (C)') +
  atltheme() 
# ggsave('R_plots/Exp042 ploidy histogram V2 no F2s.pdf', p1, device = cairo_pdf, units = 'in', width = 3.2, height = 1.9)
# ggsave('R_plots/Exp042 ploidy histogram V2 no F2s.png', p1, device = 'png', units = 'in', width = 3.2, height = 1.9, dpi = 300)
```

### sumamry statistics and hypothesis testing
```{r}
# summary stats: mean and bootstrapped 95%CI
ploidy2_summary <- ploidy2 %>%
  group_by(genotype) %>%
  do(data.frame(rbind(Hmisc::smean.cl.boot(.$c_val, B=10000)))) %>%
  mutate(genotype = factor(genotype, levels = c('balminF2', 'balmin', 'balplus')))

ggplot(ploidy2_summary, aes(genotype, color = genotype)) +
  geom_point(aes(y = Mean), size = 1.5) +
  geom_errorbar(aes(ymin = Lower, ymax = Upper), width = 0.9) +
  coord_flip() +
  scale_y_log10(breaks = c(1,2,4,8,16,32,64), limits = c(2, 40)) +
  scale_color_manual(values = c('black', '#FF4F57', '#00328A')) +
  atltheme() +
  theme(rect = element_rect(fill = "transparent"),
        panel.grid = element_blank())
ggsave('R_plots/Exp042 ploidy summary stats V2.pdf', device = cairo_pdf, units = 'in', width = 3.2485, height = 1)

p2 <- p1 + 
  geom_vline(data = mutate(ploidy2_summary, genotype = factor(genotype, levels = c('balmin', 'balminF2', 'balplus'))),
             aes(xintercept = Mean, color = genotype))
# ggsave('R_plots/Exp042 ploidy histogram V2 with means.png', p2, device = 'png', units = 'in', width = 3.2, height = 1.9, dpi = 300)

# hypothesis testing
print(kruskal.test(c_val~genotype, data = ploidy2)$p.value)
print(FSA::dunnTest(ploidy2$c_val, ploidy2$genotype, method = 'bh'))

# by worm
ploidy2_summary_byworm <- ploidy2 %>%
  group_by(genotype, image) %>%
  summarize(mean_cval = mean(c_val)) %>%
  summarize(mean = mean(mean_cval), sd = sd(mean_cval), n = n())

```

### ploidy of CDK-2 gut(-) by cell type
```{r}
ploidy2 %>%
  filter(genotype != 'balminF2') %>%
  mutate(genotype = factor(genotype, levels = c('balplus', 'balmin'), labels = c('control', 'CDK-2gut(-)'))) %>%
  ggplot(aes(c_val, fill = genotype)) +
  scale_x_log10(breaks = c(1,2,4,8,16,32,64)) +
  geom_histogram(bins = 20, position = 'identity') +
  facet_grid(cell_type~genotype) +
  scale_fill_manual(values = c('#00328A', '#FF4F57')) +
  atltheme() +
  labs(x = 'nucleus ploidy (C)') +
  theme(panel.grid.major.x = element_line(color = 'grey60', linewidth = 0.1))
ggsave('R_plots/Exp042 ploidy histogram by celltype.pdf', device = cairo_pdf, units = 'in', width = 3.9, height = 2.5)
```

## Combine with nucleus data for total ploidy graph

